Unlocking Sensitivity for Visibility-based Estimators of the 21\,cm Reionization Power Spectrum

### 2017-07-22 Rotation paper/draft_working.tex
Reconfigured equation and figure placements; fixed typos

The epoch of reionization represents the last key
stage of our Universe's early evolution. The study of this event stands at
the intersection of cosmology and astrophysics. Understanding this
event not only serves as a scientific goal
of its own, but also as a gateway to information
regarding fundamental physics of inflation, neutrino mass, and the phenomenology
of the first stars and galaxies (e.g., \citealt{LiuOpticalDepth, Liu2016b, Mao2008, DEw21cm, Bull2015, Oyama20131186}).

@@

The basic idea of our proposed visibility-based, two-baseline power-spectrum estimator is as follows. The Earth's rotation causes the baselines in a drift-scan array to pick up different modes of the sky with time. Rotation-synthesis makes use of the rotation-induced $uv$ coverage map to form an image. With visibility-based pipelines, the same information can be extracted. To do so, we cross-multiply time-shifted visibilities, with the proper weighting, to form power spectra.  Due to the large number of elements of modern arrays, the task of cross-multiplying every baseline against every other, scaling as number of array elements to the fourth power, can be computationally formidable, and many pairs of baselines provide only negligible redundancy information. Our contribution is thus twofold. First, we introduce a formalism to estimate the power spectrum from pairs of partially-redundant baselines in a visibility pipeline. Secondly we show how to use the two-baseline estimator to automatically pre-select baseline pairs and time offsets, making the problem computationally efficient. More precisely, our formalism allows one to simultaneously identify the baselines that have strong
redundancy, find the time offset that corresponds to maximal redundancy for a given pair of baselines, and quantify the sensitivity associated with cross multiplying
such a pair of partially redundant baselines, which in turn is used as weight to combine measurements in a power spectrum pipeline.

@@

We shall use the 128-element PAPER array (henceforth referred to as PAPER-128) to motivate our formalism and demonstrate our method, and extend our results to several HERA configurations in Section \ref{sec:arrconf}. 
The PAPER array is located in the Karoo desert in South Africa (30:43:17.5
S, 21:25:41.8 E). The layout pattern is shown
in Fig. \ref{fig:AntPos}. The array consists of a 112-element core in a rectangular grid, and 16 ``outriggers'' used primarily to aid calibration.  In the bottom panel, the two baselines marked in red are an example of an equivalent pair. We denote a equivalency-class of baselines in the PAPER grid by their separations, in this case  \{2,0\}, for the
antennas are separated by 2 units east and 0 units north. Similarly,
the baselines marked in blue and yellow are respectively examples
of \{2,1\} and \{2, -1\}.
Note that \{2,0\} and \{-2, 0\} are the same class of baselines and should not be counted twice. Antennas in purely north-south baselines
are close (4m), and hence these baselines are not suitable
for sensitive measurements due to cross-coupling. On the other hand, the small North-South separation means that classes such as \{2,0\} and  \{2,1\} are expected to be near-equivalent. The PAPER-64 analysis of \cite{Ali2015} uses three classes of baselines, the PAPER-128
equivalent of which are 
\{2, 0\}, \{2, 1\} and \{2, -1\}. There, baselines are cross-multiplied within each class. This paper provides the method for inter-class multiplications. We will use the short hand notation $\{m,n:m',n'\}$ to denote a pair of baseline classes to be cross-multiplied.

@@

Equivalent baselines follow identical $uvw$ tracks. Traditionally, we can identify
redundancy of nearly equivalent baselines as crossings
of the $uv$ tracks, a two-dimensional projection of $uvw$, as shown in the top panel of Fig. \ref{fig:Tracks}. However, there are several reasons that $uv$ track-crossing does not imply perfect redundancy. The most obvious reason is that the three-dimensional $uvw$ tracks do not actually cross, as is evident from the $vw$ projection in the bottom panel of Fig. \ref{fig:Tracks}. In general, track-crossings are
not accurate enough for time offset determination, nor can it give estimate of the degree of redundancy.   For drift-scan arrays, even a hypothetical crossing in $uvw$ space would not imply perfect redundancy. This will become evident in the next section, after we develop a more general formalism that accounts for the point spread function of primary. The relation between track-crossing and redundancy for both drift-scan and tracking measurements is further explored in Appendix \ref{sec:appA}.

@@

Ultimately we would like to relate the observed visibility to the power spectrum, $P$, defined such that
\[
\langle \widetilde{T}^{*}(\boldsymbol{k})\widetilde{T}(\boldsymbol{k'})\rangle = \left(2\pi\right)^3\delta_D^3(\bold{k}-\bold{k'})P(k),
\label{eq:Pofk}
\]
where $\bold{k}$ is the cosmological wavenumber, $\delta_D^3$ is the three-dimensional Dirac delta function, and $\widetilde{T}(\boldsymbol{k})$ is the three-dimensional Fourier transform of the brightness temperature field $T(\bold{r})$, with $\bold r$ being the cosmological position coordinate. 
We begin by relating the observational coordinates $\hat{\boldsymbol{s}}$
and $\nu$, to
$\boldsymbol{r}$ and $\boldsymbol{k}$: 
\[
\begin{aligned}r & =\frac{c}{H_{0}}\int_{0}^{z}\frac{dz'}{E(z')}\\
 & \approx\frac{c}{H_{0}}\int_{0}^{z_{0}}\frac{dz'}{E(z')}-\frac{c(1+z_0)^{2}}{\nu_{21}H_{0}E(z_0)}\left(\nu-\nu_{0}\right)\\
 & \equiv X-Y\Delta\nu,
\end{aligned} \label{eq:r}
\]
where $r$ is the magnitude of the vector $\bold{r}$, $\nu_{21}=1.420$\,GHz is the 21\,cm transition rest frequency, $\nu_{0}$ is
a reference central frequency with corresponding redshift $z_{0}$, $H_0$ is the current-day Hubble constant,  
and 
\[
E(z)=\sqrt{\Omega_{m}(1+z)^{3}+\Omega_{\Lambda}}.
\]
Inverting for $\nu$:
\begin{equation}
\nu=\frac{X-r}{Y}+\nu_{0}.\label{eq:nur}
\end{equation}

In the thin-shell limit, we can write:
\begin{equation}
d^2r=X^2d\Omega. 
\label{eq:thinshell}
\end{equation}
Note that Eq. \eqref{eq:thinshell} does not require the flat-sky approximation; the angular integral is still performed over the curved sky.

Eqs. \eqref{eq:nur} and \eqref{eq:thinshell} show that observational angular and spectral dimensions together correspond to the three dimensional volume of cosmological interest. Moreover, recall that the power spectrum in Eq. \eqref{eq:Pofk} is proportional to the product of Fourier transforms of the temperature field. Since Eq. \eqref{eq:Vis1} resembles a Fourier transform along the angular directions, it is natural to perform a further Fourier transform along the radial, or equivalently, frequency axis. The transform, called delay-transform \citep{PB2009}, has gained popularity recently since it has been shown that foregrounds are isolated in delay space\footnote{We discuss foreground isolation in detail in Section \ref{sec:chromaticity}.}. We define the delay-transformed visibility as:
\small
\begin{equation}
\begin{aligned}V(\boldsymbol{b},\tau) & =\int d\nu V_{\nu}(\boldsymbol{b})\phi(\nu)\exp\left(2\pi i\nu\tau\right)\\
 & =\int d\Omega d\nu B(\hat{\boldsymbol{s}},\nu)T(\hat{\boldsymbol{s}},\nu)\exp\left[-2\pi i\nu\left(\frac{\boldsymbol{b}\cdot\hat{\boldsymbol{s}}}{c}-\tau\right)\right]. 
\end{aligned}
\label{eq:Vb1}
\end{equation}
\normalsize
Here the delay $\tau$ is the Fourier dual of $\nu$. Eq. \eqref{eq:Vb1} expresses the delay-transformed visibility as
an integral over observational coordinates $\hat{\boldsymbol{s}}$ and $\nu$. For notational simplicity we have defined the quantity

@@

With Eqs. \eqref{eq:nur} and \eqref{eq:thinshell}, we can rewrite the delayed-transformed visibility in cosmological coordinates as 
\small
\[
\begin{aligned}V(\boldsymbol{b},\tau) & =\int\frac{d^{3}r}{X^{2}Y}B(\boldsymbol{r})T(\boldsymbol{r})\exp\left[-2\pi i\left(\frac{\boldsymbol{b}}{c}\cdot\hat{\boldsymbol{r}}-\tau\right)\nu_{r}\right],
\end{aligned}
\]
\normalsize
where $d\nu=-dr/Y$ and $d^{3}r=-X^{2}Yd\Omega d\nu$. 
We have written $\nu_{r}$ as a reminder that $\nu$ and $r$ are related
by Eq. \eqref{eq:nur}.

Existing visibility-based power-spectrum pipelines for redundant drift-scan  arrays relate the power-spectrum to the conjugate square of the visibilities \citep{delay-transform, paper32, Ali2015}. We would like to generalize such relations by relating the power-spectrum to the product of two visibilities from two arbitrary baselines and time offsets. 
To do this, we must account for the fact that the beam pattern of a baseline shifts relative to the sky as the Earth rotates. Here we choose to fix the sky, and denote the rotated coordinates
in the topocentric frame with the three-dimensional rotation operator $\boldsymbol{\Gamma}$:

\[
\begin{aligned}V_{\psi}(\boldsymbol{b'},\tau) & = \int\frac{d^{3}r}{X^{2}Y}B(\boldsymbol{\Gamma}\boldsymbol{r})T(\boldsymbol{r})\\
& \qquad \qquad  \exp\left[-2\pi i\left(\frac{\boldsymbol{b}}{c}\cdot\boldsymbol{\Gamma}\hat{\boldsymbol{r}}-\tau\right)\nu_{r}-i\psi_\nu\right],\end{aligned}
\]
where we introduced a frequency-dependent phase $\psi_{\nu}$, which to linear-order in $\nu$ corresponds to phasing the two visibilities to the same phase center. \footnote{For many cases, the linear-order interpretation is sufficient. We keep the general term to reserve the option to determine its exact form numerically as this information comes at no additional computational cost. }

With implicit bounds of integrals from $-\infty$ to $\infty$, we have:
\begin{widetext}
\begin{equation}
\begin{aligned} \langle V^{*}(\boldsymbol{b},\tau)V_{\psi}(\boldsymbol{b'},\tau)\rangle
 & =\int \frac{d^{3}rd^{3}r'}{(X^2Y)^2}\langle T^{*}(\boldsymbol{r})T(\boldsymbol{r'})\rangle B^{*}(\boldsymbol{r})B(\boldsymbol{\Gamma} \boldsymbol{r'})\Phi(\boldsymbol{r},\boldsymbol{\Gamma} \boldsymbol{r'})\\
 & =\int \frac{d^{3}rd^{3}r'}{(X^2Y)^2}\left(\int\frac{d^{3}\kappa}{(2\pi)^{3}}\frac{d^{3}\kappa'}{(2\pi)^{3}}\langle T^{*}(\boldsymbol{\kappa})T(\boldsymbol{\kappa'})\rangle e^{-i(\boldsymbol{\kappa}\cdot \boldsymbol{r}-\boldsymbol{\kappa'}\cdot\boldsymbol{r'})}\right)B^{*}(\boldsymbol{r})B(\boldsymbol{\Gamma} \boldsymbol{r'})\Phi(\boldsymbol{r},\boldsymbol{\Gamma} \boldsymbol{r'})\\
 & =\int \frac{d^{3}rd^{3}r'}{(X^2Y)^2}\left(\int\frac{d^{3}\kappa}{(2\pi)^{3}}P(\kappa)e^{-i\boldsymbol{\kappa}\cdot(\boldsymbol{r}-\boldsymbol{r'})}\right)B^{*}(\boldsymbol{r})B(\boldsymbol{\Gamma} \boldsymbol{r'})\Phi(\boldsymbol{r},\boldsymbol{\Gamma} \boldsymbol{r'})\\
 & \approx P(k_{\bar{b}, \tau})\int \frac{d^{3}rd^{3}r'}{(X^2Y)^2} \delta_{D}^{(3)}(\boldsymbol{r}-\boldsymbol{r'})B^{*}(\boldsymbol{r})B(\boldsymbol{\Gamma} \boldsymbol{r'}) \Phi(\boldsymbol{r},\boldsymbol{\Gamma} \boldsymbol{r'})\\
 & = P(k_{\bar{b}, \tau})\int \frac{d^{3}r}{(X^2Y)^2}B^{*}(\boldsymbol{r})B(\boldsymbol{\Gamma} \boldsymbol{r})\exp\left[-i2\pi\frac{\nu_{r}}{c}\left(\hat{\boldsymbol{r}}\cdot\boldsymbol{b}-\boldsymbol{\Gamma} \hat{\boldsymbol{r}}\cdot\boldsymbol{b'}\right)-i\psi_{\nu}\right],
\end{aligned}
\label{eq:main}
\end{equation}

where $\delta_D$ is the Dirac delta-function and
\begin{equation}
\Phi(\boldsymbol{r},\boldsymbol{\Gamma} \boldsymbol{r'})\equiv\exp\left[i\frac{2\pi}{c}\left(\boldsymbol{b}\cdot\nu_{r}\hat{\boldsymbol{r}}-\boldsymbol{b'}\cdot\nu_{r'}\boldsymbol{\Gamma}\hat{\boldsymbol{r}}'\right)\right]\exp\left[-i2\pi\tau\left(\nu_{r}-\nu_{r'}\right)-i\psi_{\nu}\right].
\end{equation}

\end{widetext}

The third equality of Eq.(\ref{eq:main}) follows from assuming the translational invariance of statistics of the 21\,cm field. 
For the inequality in the fourth line, we have assumed that the  3D power spectrum varies negligibly over the $k$-space of interest and have thus pulled out of the integral the power spectrum centered at 
\[
k_{\boldsymbol{b}, \tau} \equiv 2\pi \sqrt{\left(\frac{\tau}{Y}\right)^2 + \left(\frac{\bar{b}}{\lambda X}\right)^2},
\label{eq:kbtau}
\]
where $\bar{b}$ is approximately the mean of $b$ and $b'$. 
In Section \ref{sec:chromaticity} we will examine the exact form of the Fourier-space footprint probed as well as its effect on foreground isolation.  
%Since $\boldsymbol{\Gamma}$ is a sky rotation, it does not affect $\nu$, hence we have taken $\nu_{r}$ outside the parenthesis in the phase term $\exp\left[-i2\pi\frac{\nu_{r}}{c}\left(\hat{\boldsymbol{r}}\cdot\boldsymbol{b}-\boldsymbol{\Gamma} \hat{\boldsymbol{r}}\cdot\boldsymbol{b'}\right)\right]$. 
Notice that the phase factor $\exp\left[-i2\pi\tau\left(\nu-\nu'\right)\right]$
drops out in the end.

Since the beam pattern and bandpass are given in $\hat{\boldsymbol{s}}$
and $\nu$, we convert the last line of Eq. \eqref{eq:main} back to these coordinates to get
the general relation between the delay-transformed visibilities and
the power spectrum:

\begin{equation}
\begin{aligned} & \langle V^{*}(\boldsymbol{b},\tau)V_{\psi}(\boldsymbol{b'},\tau)\rangle\\
 & \qquad \approx P(k_{\bar{b}, \tau})\int\frac{d\Omega d\nu}{X^{2}Y}B^{*}(\hat{\boldsymbol{s}},\nu)B(\boldsymbol{\Gamma}\hat{\boldsymbol{s}},\nu)\\
 & \qquad \qquad \qquad \qquad \exp\left[i2\pi\frac{\nu}{c}\left(\hat{\boldsymbol{s}}\cdot\boldsymbol{b}-\boldsymbol{\Gamma}\hat{\boldsymbol{s}}\cdot\boldsymbol{b'}\right)-i\psi_{\nu}\right].
 \end{aligned}
\label{eq:final}
\end{equation}

We can therefore form the power spectrum estimator for the baseline pair $\{\boldsymbol b: \boldsymbol  b'\}$:

\begin{equation}
 \hat{P}(k_{\bar{b}, \tau}) \equiv \frac{V^{*}(\boldsymbol{b},\tau)V_{\psi}(\boldsymbol{b'},\tau)}{\Theta}, 
 \label{eq:opp}
\end{equation}
where the weight is defined as 
\begin{equation}
\Theta \equiv\int d\nu \Theta_{\nu}, 
\label{eq:Theta}
\end{equation}
with

\[
\Theta_{\nu} \equiv e^{-i\psi_{\nu}} \int \frac{d\Omega}{X^{2}Y}B^{*}(\hat{\boldsymbol{s}},\nu)B(\boldsymbol{\Gamma}\hat{\boldsymbol{s}},\nu) e^{i2\pi\frac{\nu}{c}\left(\hat{\boldsymbol{s}}\cdot\boldsymbol{b}-\boldsymbol{\Gamma}\hat{\boldsymbol{s}}\cdot\boldsymbol{b'}\right)}.
\label{eq:Thetanu}
\]

Notice that $\Theta$ has no dependence on $\tau$. We point out that although all our derivations focused on drift-scan telescopes, we can get the analogous result for tracking measurements simply by noticing that for a tracking primary beam with radial symmetry we have:
\[
\Theta_{\nu} \equiv e^{-i\psi_{\nu}} \int \frac{d\Omega}{X^{2}Y}B^{*}(\hat{\boldsymbol{s}},\nu)B(\hat{\boldsymbol{s}},\nu) e^{i2\pi\frac{\nu}{c}\left(\hat{\boldsymbol{s}}\cdot\boldsymbol{b}-\boldsymbol{\Gamma}\hat{\boldsymbol{s}}\cdot\boldsymbol{b'}\right)}.
\label{eq:Thetanu_tracking}
\]

>>>>

The Epoch of Reionization (EoR) represents the last key
stage of our Universe's early evolution. The study of this event stands at
the intersection of cosmology and astrophysics. Understanding this
event not only serves as a scientific goal
of its own, but also as a gateway to information
regarding fundamental physics of inflation, neutrino mass, and the phenomenology
of the first stars and galaxies (e.g., \citealt{LiuOpticalDepth, Liu2016b, Mao2008, DEw21cm, Bull2015, Oyama20131186}).

@@

The basic idea of our proposed visibility-based, two-baseline power-spectrum estimator is as follows. The Earth's rotation causes the baselines in a drift-scan array to pick up different modes of the sky with time. Rotation synthesis makes use of the rotation-induced $uv$ coverage map to form an image. With visibility-based pipelines, the same information can be extracted. To do so, we cross-multiply time-shifted visibilities, with the proper weighting, to form power spectra.  Due to the large number of elements of modern arrays, the task of cross-multiplying every baseline against every other, scaling as number of array elements to the fourth power, can be computationally formidable, and many pairs of baselines provide only negligible redundancy information. Our contribution is thus twofold. First, we introduce a formalism to estimate the power spectrum from pairs of partially-redundant baselines in a visibility pipeline. Secondly we show how to use the two-baseline estimator to automatically pre-select baseline pairs and time offsets, making the problem computationally efficient. More precisely, our formalism allows one to simultaneously identify the baselines that have strong
redundancy, find the time offset that corresponds to maximal redundancy for a given pair of baselines, and quantify the sensitivity associated with cross multiplying
such a pair of partially redundant baselines, which in turn is used as weight to combine measurements in a power spectrum pipeline.

@@

We shall use the 128-element PAPER array (henceforth referred to as PAPER-128) to motivate our formalism and demonstrate our method, and extend our results to several HERA configurations in Section \ref{sec:arrconf}. 
The PAPER array is located in the Karoo desert in South Africa (30:43:17.5
S, 21:25:41.8 E). The layout pattern is shown
in Fig. \ref{fig:AntPos}. The array consists of a 112-element core in a rectangular grid, and 16 ``outriggers'' used primarily to aid calibration.  In the bottom panel, the two baselines marked in red are an example of an equivalent pair. We denote a equivalency class of baselines in the PAPER grid by their separations, in this case  \{2,0\}, for the
antennas are separated by 2 units east and 0 units north. Similarly,
the baselines marked in blue and yellow are respectively examples
of \{2,1\} and $\{2, -1\}$.
Note that \{2,0\} and $\{-2, 0\}$ are the same class of baselines and should not be counted twice. Antennas in purely north-south baselines
are close (4m), and hence these baselines are not suitable
for sensitive measurements due to cross-coupling. On the other hand, the small North-South separation means that classes such as \{2,0\} and  \{2,1\} are expected to be near-equivalent. The PAPER-64 analysis of \cite{Ali2015} uses three classes of baselines, the PAPER-128
equivalent of which are 
\{2, 0\}, \{2, 1\} and $\{2, -1\}$. There, baselines are cross-multiplied within each class. This paper provides the method for inter-class multiplications. We will use the short hand notation $\{m,n:m',n'\}$ to denote a pair of baseline classes to be cross-multiplied.

@@

Equivalent baselines follow identical $uvw$ tracks. Traditionally, we can identify
redundancy of nearly equivalent baselines as crossings
of the $uv$ tracks, a two-dimensional projection of $uvw$, as shown in the top panel of Fig. \ref{fig:Tracks}. However, there are several reasons that $uv$ track-crossing does not imply perfect redundancy. The most obvious reason is that the three-dimensional $uvw$ tracks do not actually cross, as is evident from the $vw$ projection in the bottom panel of Fig. \ref{fig:Tracks}. In general, track-crossings are
not accurate enough to determine the optimal time offset for the cross multiplication of visibilities, nor can it provide an estimate of the degree of redundancy.   For drift-scan arrays, even a hypothetical crossing in $uvw$ space would not imply perfect redundancy. This will become evident in the next section, after we develop a more general formalism that accounts for the point spread function of primary. The relation between track-crossing and redundancy for both drift-scan and tracking measurements is further explored in Appendix \ref{sec:appA}.

@@

Ultimately we would like to relate the observed visibility to the power spectrum, $P$, defined such that
\[
\langle \widetilde{T}^{*}(\boldsymbol{k})\widetilde{T}(\boldsymbol{k'})\rangle = \left(2\pi\right)^3\delta_D^3(\bold{k}-\bold{k'})P(k),
\label{eq:Pofk}
\]
where $\bold{k}$ is the cosmological wavenumber, $\delta_D^3$ is the three-dimensional Dirac delta function, and $\widetilde{T}(\boldsymbol{k})$ is the three-dimensional Fourier transform of the brightness temperature field $T(\bold{r})$, with $\bold r$ being the cosmological position coordinate. 
We begin by relating the observational coordinates $\hat{\boldsymbol{s}}$
and $\nu$, to
$\boldsymbol{r}$ and $\boldsymbol{k}$: 
\[
\begin{aligned}r & =\frac{c}{H_{0}}\int_{0}^{z}\frac{dz'}{E(z')}\\
 & \approx\frac{c}{H_{0}}\int_{0}^{z_{0}}\frac{dz'}{E(z')}-\frac{c(1+z_0)^{2}}{\nu_{21}H_{0}E(z_0)}\left(\nu-\nu_{0}\right)\\
 & \equiv X-Y\Delta\nu,
\end{aligned} \label{eq:r}
\]
where $r$ is the magnitude of the vector $\bold{r}$, $\nu_{21}=1.420$\,GHz is the 21\,cm transition rest frequency, $\nu_{0}$ is
a reference central frequency with corresponding redshift $z_{0}$, $H_0$ is the current-day Hubble constant,  
and 
\[
E(z)=\sqrt{\Omega_{m}(1+z)^{3}+\Omega_{\Lambda}}.
\]
Inverting for $\nu$ gives
\begin{equation}
\nu=\frac{X-r}{Y}+\nu_{0}.\label{eq:nur}
\end{equation}
This provides us with a mapping between the observational frequency and the cosmological distance in the radial direction. For the angular directions, we have in the thin-shell limit
\begin{equation}
d^2r=X^2d\Omega. 
\label{eq:thinshell}
\end{equation}
Note that Eq. \eqref{eq:thinshell} does not require the flat-sky approximation; the angular integral is still performed over the curved sky.

Eqs. \eqref{eq:nur} and \eqref{eq:thinshell} show that observational angular and spectral dimensions together correspond to the three dimensional volume of cosmological interest. Moreover, recall that the power spectrum in Eq. \eqref{eq:Pofk} is proportional to the product of Fourier transforms of the temperature field. Since Eq. \eqref{eq:Vis1} resembles a Fourier transform along the angular directions, it is natural to perform a further Fourier transform along the radial, or equivalently, frequency axis. The transform, called the delay transform \citep{PB2009}, has gained popularity recently since it has been shown that foregrounds are isolated in delay space.\footnote{We discuss foreground isolation in detail in Section \ref{sec:chromaticity}.} We define the delay-transformed visibility as:
\small
\begin{equation}
\begin{aligned}V(\boldsymbol{b},\tau) & \equiv \int d\nu V_{\nu}(\boldsymbol{b})\phi(\nu)\exp\left(2\pi i\nu\tau\right)\\
 & =\int d\Omega d\nu B(\hat{\boldsymbol{s}},\nu)T(\hat{\boldsymbol{s}},\nu)\exp\left[-2\pi i\nu\left(\frac{\boldsymbol{b}\cdot\hat{\boldsymbol{s}}}{c}-\tau\right)\right]. 
\end{aligned}
\label{eq:Vb1}
\end{equation}
\normalsize
Here the delay $\tau$ is the Fourier dual of $\nu$. Eq. \eqref{eq:Vb1} expresses the delay-transformed visibility as
an integral over observational coordinates $\hat{\boldsymbol{s}}$ and $\nu$. For notational simplicity we have defined the quantity

@@

With Eqs. \eqref{eq:nur} and \eqref{eq:thinshell}, we can rewrite the delayed-transformed visibility in cosmological coordinates as 
\small
\[
\begin{aligned}V(\boldsymbol{b},\tau) & =\int\frac{d^{3}r}{X^{2}Y}B(\boldsymbol{r})T(\boldsymbol{r})\exp\left[-2\pi i\left(\frac{\boldsymbol{b}}{c}\cdot\hat{\boldsymbol{r}}-\tau\right)\nu_{r}\right],
\end{aligned}
\]
\normalsize
\!\!where $d\nu=-dr/Y$ and $d^{3}r=-X^{2}Yd\Omega d\nu$. 
We have written $\nu_{r}$ as a reminder that $\nu$ and $r$ are related
by Eq. \eqref{eq:nur}.

Existing visibility-based power-spectrum pipelines for redundant drift-scan  arrays relate the power-spectrum to the conjugate square of the visibilities \citep{delay-transform, paper32, Ali2015}. We would like to generalize such relations by relating the power-spectrum to the product of two visibilities from two arbitrary baselines and time offsets. 
To do this, we must account for the fact that the beam pattern of a baseline shifts relative to the sky as the Earth rotates. Here we choose to fix the sky, and denote the rotated coordinates
in the topocentric frame with the three-dimensional rotation operator $\boldsymbol{\Gamma}$:
\[
\begin{aligned}V_{\psi}(\boldsymbol{b'},\tau) & = \int\frac{d^{3}r}{X^{2}Y}B(\boldsymbol{\Gamma}\boldsymbol{r})T(\boldsymbol{r}) e^{-2\pi i\left(\frac{\boldsymbol{b}}{c}\cdot\boldsymbol{\Gamma}\hat{\boldsymbol{r}}-\tau\right)\nu_{r}-i\psi_\nu},\end{aligned}
\]
%\[
%\begin{aligned}V_{\psi}(\boldsymbol{b'},\tau) & = \int\frac{d^{3}r}{X^{2}Y}B(\boldsymbol{\Gamma}\boldsymbol{r})T(\boldsymbol{r}) \times\\
%& \qquad \qquad  \exp\left[-2\pi i\left(\frac{\boldsymbol{b}}{c}\cdot\boldsymbol{\Gamma}\hat{\boldsymbol{r}}-\tau\right)\nu_{r}-i\psi_\nu\right],\end{aligned}
%\]
where we introduced a frequency-dependent phase $\psi_{\nu}$, which to linear-order in $\nu$ one would typically pick to correspond to the re-phasing of the two visibilities to the same phase center.\footnote{For many cases, the linear-order interpretation is sufficient. We keep the general term to reserve the option to determine its exact form numerically as this information comes at no additional computational cost.}

With implicit bounds of integrals from $-\infty$ to $\infty$, we have:
\begin{eqnarray}
 &&\langle V^{*}(\boldsymbol{b},\tau)V_{\psi}(\boldsymbol{b'},\tau)\rangle \nonumber \\
 && =\int \frac{d^{3}rd^{3}r'}{(X^2Y)^2}\langle T^{*}(\boldsymbol{r})T(\boldsymbol{r'})\rangle B^{*}(\boldsymbol{r})B(\boldsymbol{\Gamma} \boldsymbol{r'})\Phi(\boldsymbol{r},\boldsymbol{\Gamma} \boldsymbol{r'}) \nonumber \\
 &&  =\int \frac{d^{3}rd^{3}r'}{(X^2Y)^2}\left(\int\frac{d^{3}\kappa}{(2\pi)^{3}}\frac{d^{3}\kappa'}{(2\pi)^{3}}\langle T^{*}(\boldsymbol{\kappa})T(\boldsymbol{\kappa'})\rangle e^{-i(\boldsymbol{\kappa}\cdot \boldsymbol{r}-\boldsymbol{\kappa'}\cdot\boldsymbol{r'})}\right) \nonumber \\
 && \qquad  \times B^{*}(\boldsymbol{r})B(\boldsymbol{\Gamma} \boldsymbol{r'})\Phi(\boldsymbol{r},\boldsymbol{\Gamma} \boldsymbol{r'})\nonumber \\
 && =\int \frac{d^{3}rd^{3}r'}{(X^2Y)^2}\left(\int\frac{d^{3}\kappa}{(2\pi)^{3}}P(\kappa)e^{-i\boldsymbol{\kappa}\cdot(\boldsymbol{r}-\boldsymbol{r'})}\right) \nonumber \\
 && \qquad \times B^{*}(\boldsymbol{r})B(\boldsymbol{\Gamma} \boldsymbol{r'})\Phi(\boldsymbol{r},\boldsymbol{\Gamma} \boldsymbol{r'}) \nonumber \\
 && \approx P(k_{\bar{b}, \tau})\int \frac{d^{3}rd^{3}r'}{(X^2Y)^2} \delta_{D}^{(3)}(\boldsymbol{r}-\boldsymbol{r'})B^{*}(\boldsymbol{r})B(\boldsymbol{\Gamma} \boldsymbol{r'}) \Phi(\boldsymbol{r},\boldsymbol{\Gamma} \boldsymbol{r'}) \nonumber \\
 && = P(k_{\bar{b}, \tau})\int \frac{d^{3}r}{(X^2Y)^2}B^{*}(\boldsymbol{r})B(\boldsymbol{\Gamma} \boldsymbol{r}) e^{-i2\pi\frac{\nu_{r}}{c}\left(\hat{\boldsymbol{r}}\cdot\boldsymbol{b}-\boldsymbol{\Gamma} \hat{\boldsymbol{r}}\cdot\boldsymbol{b'}\right)-i\psi_{\nu}}, \qquad
\label{eq:main}
\end{eqnarray}
where $\delta_D$ is the Dirac delta-function and
\begin{eqnarray}
\Phi(\boldsymbol{r},\boldsymbol{\Gamma} \boldsymbol{r'})\equiv &\exp\left[i\frac{2\pi}{c}\left(\boldsymbol{b}\cdot\nu_{r}\hat{\boldsymbol{r}}-\boldsymbol{b'}\cdot\nu_{r'}\boldsymbol{\Gamma}\hat{\boldsymbol{r}}'\right)\right] \nonumber \\
& \times \exp\left[-i2\pi\tau\left(\nu_{r}-\nu_{r'}\right)-i\psi_{\nu}\right].
\end{eqnarray}
%
%\begin{widetext}
%\begin{equation}
%\begin{aligned} \langle V^{*}(\boldsymbol{b},\tau)V_{\psi}(\boldsymbol{b'},\tau)\rangle
% & =\int \frac{d^{3}rd^{3}r'}{(X^2Y)^2}\langle T^{*}(\boldsymbol{r})T(\boldsymbol{r'})\rangle B^{*}(\boldsymbol{r})B(\boldsymbol{\Gamma} \boldsymbol{r'})\Phi(\boldsymbol{r},\boldsymbol{\Gamma} \boldsymbol{r'})\\
% & =\int \frac{d^{3}rd^{3}r'}{(X^2Y)^2}\left(\int\frac{d^{3}\kappa}{(2\pi)^{3}}\frac{d^{3}\kappa'}{(2\pi)^{3}}\langle T^{*}(\boldsymbol{\kappa})T(\boldsymbol{\kappa'})\rangle e^{-i(\boldsymbol{\kappa}\cdot \boldsymbol{r}-\boldsymbol{\kappa'}\cdot\boldsymbol{r'})}\right)B^{*}(\boldsymbol{r})B(\boldsymbol{\Gamma} \boldsymbol{r'})\Phi(\boldsymbol{r},\boldsymbol{\Gamma} \boldsymbol{r'})\\
% & =\int \frac{d^{3}rd^{3}r'}{(X^2Y)^2}\left(\int\frac{d^{3}\kappa}{(2\pi)^{3}}P(\kappa)e^{-i\boldsymbol{\kappa}\cdot(\boldsymbol{r}-\boldsymbol{r'})}\right)B^{*}(\boldsymbol{r})B(\boldsymbol{\Gamma} \boldsymbol{r'})\Phi(\boldsymbol{r},\boldsymbol{\Gamma} \boldsymbol{r'})\\
% & \approx P(k_{\bar{b}, \tau})\int \frac{d^{3}rd^{3}r'}{(X^2Y)^2} \delta_{D}^{(3)}(\boldsymbol{r}-\boldsymbol{r'})B^{*}(\boldsymbol{r})B(\boldsymbol{\Gamma} \boldsymbol{r'}) \Phi(\boldsymbol{r},\boldsymbol{\Gamma} \boldsymbol{r'})\\
% & = P(k_{\bar{b}, \tau})\int \frac{d^{3}r}{(X^2Y)^2}B^{*}(\boldsymbol{r})B(\boldsymbol{\Gamma} \boldsymbol{r})\exp\left[-i2\pi\frac{\nu_{r}}{c}\left(\hat{\boldsymbol{r}}\cdot\boldsymbol{b}-\boldsymbol{\Gamma} \hat{\boldsymbol{r}}\cdot\boldsymbol{b'}\right)-i\psi_{\nu}\right],
%\end{aligned}
%\label{eq:main}
%\end{equation}
%
%where $\delta_D$ is the Dirac delta-function and
%\begin{equation}
%\Phi(\boldsymbol{r},\boldsymbol{\Gamma} \boldsymbol{r'})\equiv\exp\left[i\frac{2\pi}{c}\left(\boldsymbol{b}\cdot\nu_{r}\hat{\boldsymbol{r}}-\boldsymbol{b'}\cdot\nu_{r'}\boldsymbol{\Gamma}\hat{\boldsymbol{r}}'\right)\right]\exp\left[-i2\pi\tau\left(\nu_{r}-\nu_{r'}\right)-i\psi_{\nu}\right].
%\end{equation}
%
%\end{widetext}
The third equality of Eq.(\ref{eq:main}) follows from assuming the translational invariance of the statistics of the 21\,cm field. 
For the inequality in the fourth line, we have assumed that the  3D power spectrum varies negligibly over the $k$-space of interest and have thus pulled out of the integral the power spectrum centered at 
\[
k_{\boldsymbol{b}, \tau} \equiv 2\pi \sqrt{\left(\frac{\tau}{Y}\right)^2 + \left(\frac{\bar{b}}{\lambda X}\right)^2},
\label{eq:kbtau}
\]
where $\bar{b}$ is approximately the mean of $b$ and $b'$. 
In Section \ref{sec:chromaticity} we will examine the exact form of the Fourier-space footprint that is being probed, as well as its effect on foreground isolation.  
%Since $\boldsymbol{\Gamma}$ is a sky rotation, it does not affect $\nu$, hence we have taken $\nu_{r}$ outside the parenthesis in the phase term $\exp\left[-i2\pi\frac{\nu_{r}}{c}\left(\hat{\boldsymbol{r}}\cdot\boldsymbol{b}-\boldsymbol{\Gamma} \hat{\boldsymbol{r}}\cdot\boldsymbol{b'}\right)\right]$. 
Notice that the phase factor $\exp\left[-i2\pi\tau\left(\nu-\nu'\right)\right]$
drops out in the end.
